Shallow EDSLs and Object-Oriented Programming

### 2018-09-18 Programming19/Main.tex
Abstract

{\bf Inquiry.} Shallow embeddings are quite simple, 
but they have been criticized in the past for being quite limited in
terms of modularity and reuse. In particular it is often argued that 
shallow embeddings can only support a single DSL interpretation.

{\bf Approach.}
In this paper we argue that 
shallow embedded domain-specific languages (EDSLs) and Object-Oriented
Programming (OOP) are closely related. Gibbons and Wu already
discussed the relationship between shallow EDSLs and procedural
abstraction, while Cook discussed the connection between procedural
abstraction and OOP. We make the transitive step in this paper by
connecting shallow EDSLs directly to OOP via procedural abstraction. 
The knowledge about this relationship enables us to
improve on implementation techniques for EDSLs.

{\bf Knowledge.}
This pearl argues that common OOP mechanisms
(including \emph{inheritance}, \emph{subtyping}, and
\emph{type-refinement})
increase the modularity and reuse of shallow
EDSLs when compared to classical procedural abstraction by enabling a
simple way to express \emph{multiple, possibly dependent,
  interpretations}.

{\bf Grounding}
We make our
arguments by taking a recent paper by Gibbons and Wu, where procedural
abstraction is used in Haskell to model a simple shallow EDSL, and we recode
that EDSL both in Scala and with an improved OO-inspired Haskell encoding. 
We further illustrate our approach with
a case study on refactoring a deep external SQL DSL implementation to make it
more modular, shallow, and embedded.

{\bf Importance.} This work is important for two different
reasons. Firstly, from an intelectual point of view, this work
establishes the connection between shallow embeddings and 
Object-Oriented Programming, which enables a better understanding
of both concepts. Secondly, this work
offers new programming techniques that can be used to improve 
the modularity and reuse of shallow EDSLs.  
\end{abstract}

>>>>

{\bf Inquiry.} Shallow embeddings are quite simple, 
but they have been criticized in the past for being quite limited in
terms of modularity and reuse. In particular it is often argued that 
supporting multiple DSL interpretations in shallow embeddings is difficult.

{\bf Approach.}
This paper argues that shallow EDSLs and Object-Oriented
Programming (OOP) are closely related. Gibbons and Wu already
discussed the relationship between shallow EDSLs and procedural
abstraction, while Cook discussed the connection between procedural
abstraction and OOP. We make the transitive step in this paper by
connecting shallow EDSLs directly to OOP via procedural abstraction. 
The knowledge about this relationship enables us to
improve on implementation techniques for EDSLs.

{\bf Knowledge.}
This paper argues that common OOP mechanisms
(including \emph{inheritance}, \emph{subtyping}, and
\emph{type-refinement})
increase the modularity and reuse of shallow
EDSLs when compared to classical procedural abstraction by enabling a
simple way to express \emph{multiple, possibly dependent,
  interpretations}.

{\bf Grounding}
We make our arguments by using Gibbons and Wu's examples, where procedural
abstraction is used in Haskell to model a simple shallow EDSL. We recode that
EDSL in Scala and with an improved OO-inspired Haskell encoding.
We further illustrate our approach with
a case study on refactoring a deep external SQL DSL implementation to make it
more modular, shallow, and embedded.

{\bf Importance.} This work is important for two different
reasons. Firstly, from an intelectual point of view, this work
establishes the connection between shallow embeddings and OOP, which enables a
better understanding of both concepts. Secondly, this work
offers new programming techniques that can be used to improve 
the modularity and reuse of shallow EDSLs.  
\end{abstract}
